Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2019-05-15 IEEE_ConferenceOnGames/new_article.tex
first part of approach

@@

    At his level the composite generator has the following steps, for more details refer to the paper:

@@

    The Angry Birds level generation track is a competition where the objective is 
    to generate levels that are \textit{fun} and \textit{enjoyable} however they 
    also have to be challenging but not imposible. However the rules for the competition 
    specify that a configuration file must be used to
    generate the levels of the game in a limited amount of time, the information on
    the configuration file is first the number of \textbf{\textit{different}} levels
    that must be generated, the second value is the different materials \textit{not
    allowed} to be used, for instance a group of generated levels must not have
    circles which composition material is ice or stone, moreover it does not have to
    integrate a triangle piece with ice, stone or wood material so this combinations
    must be removed from the pool of options, the third configuration value is the
    \textit{number of pigs} that must be placed in a level, and finally the
    \textit{limit} of time that the system has to create the required amount of
    levels.

    After taking into account this restrictions the structure of our current approach
    had to be changed, so this section will be divided in two subsections, the first
    will explain the approach that was used and after that the second will explain the
    modifications that had to be made in order to integrate the configuration files
    for the competition.

    \subsection{Original approach}
    % This sentence needs more work -MG

    During the original phase of the design of the system, it was decided that the way 
    composites were going to evolve was by first placing only base pieces on a level 
    then after a simulation the remaining pieces
    of a member of the population were going to be used as a new composite if after a 
    simulation a group was still standing and grouped on a single point, their individual 
    offsets from the center of the group was calculated and then using this information a new 
    composite could be added to the pool of options, this new
    % we don't use member of the populations, use proposed solution, candidate solution
    % or individual of the population. - MG  
    additions to the composite pool could then be added to the population by the mutation
    operation when a new member of the population was being created, the objective 
    in this first phase of the project was to generate
    structures that were capable of maintaining their own balance as shown on figure
    \ref{test_old}, since the objective of open-ended evolution for this project is
    to add new composites to the pool of options, then the algorithm will first run a
    simulation with a group of pieces and after it the ones that were not destroyed
    by falling because of gravity will be saved and added to the pool of options.

@@

    Since there has to be a way to represent the structure of a level in a a way that
    both the algorithm as well as the game are able to interpret it \cite{togelius2016Representationsforsearch-basedmethods}, 
    then the way to represent the elements of a member of the population in thos work 
    is by using the id of each piece and the new added ones has a pointer to the respective data
    that needs to be obtained, this way and individual can be represented as shown
    on figure \ref{old_chrom} where each chromosome represents a certain item in the
    pool of pieces, the numbers for this elements are assigned in a first come first
    served basis, the numbers are infinitely incremented and the same pieces can be
    repeated on each individual of the population.

@@

    Once the pieces have been added and the individuals are ready to be evaluated a
    simulation on the game environment is scheduled by order in the population, all
    the individuals have a maximum of ten seconds to maintain balance or reach
    absolute stability, if absolute stability is reached it means one of two things,
    first the generated structure is completely stable or second, the tower fell to
    either side and the pieces were destroyed or stopped moving by gravitational
    pull, when either of this happens by more than two seconds the simulation ends
    regardless of time remaining and the next individual is simulated.

    The evaluation process for this project is still in the development stages,
    however in order to check that the evolution is working has it should a criteria
    has been temporarily made as follows, each individual in the population has 100
    points so before the simulation starts each of them are viable candidates to be
    an elite member of the population regardless relative height or complexity, then
    the remaining of the process is separated in three phases has follows.

@@

    Second, after removing percentage according to the missing pieces an individual
    remaining pieces are checked for their positions in comparison to the original
    ones before the simulation, using the formula 1 each piece within the individual
    is evaluated according to the center point of the piece, if the piece moved more
    than a certain threshold half the percentage for one piece is removed from the
    remaining percentage, after this using formula 2 the position of the piece
    is further checked, this time by the difference in angle from the original one,
    in case the difference of angles is more than a certain threshold half a piece
    value is further removed from the total.

>>>>

    \subsection{Composite Generation}

@@

    At this level the composite generator has the following steps:

@@

    The Angry Birds level generation track is a competition where the objective is 
    to generate levels that are \textit{fun} and \textit{enjoyable} however they 
    also have to be challenging but not imposible. However the rules for the competition 
    specify that a configuration file must be used to
    generate the levels of the game in a limited amount of time, the information on
    the configuration file is first the number of \textit{different} levels
    that must be generated, the second value is the different materials \textit{not
    allowed} to be used, for instance a group of generated levels must not have
    circles which composition material is ice or stone, moreover it does not have to
    integrate a triangle piece with ice, stone or wood material so this combinations
    must be removed from the pool of options, the third configuration value is the
    \textit{number of pigs} that must be placed in a level, and finally the
    \textit{limit} of time that the system has to create the required amount of
    levels.

    %After taking into account this restrictions the structure of our current approach
    %had to be changed, so this section will be divided in two subsections, the first
    %will explain the approach that was used and after that the second will explain the
    %modifications that had to be made in order to integrate the configuration files
    %for the competition.

    In our original approach, composites could added to the structure generator
    algorithm at runtime as a mutation of an individual. The structure generator
    is explained in the next section.

    % Lets check this, I think is not needed. 
    %The objective in this first phase of the project was to
    %generate structures that were capable of maintaining their own balance as
    %shown on figure \ref{test_old}, since the objective of open-ended evolution
    %for this project is to add new composites to the pool of options,

    In our current approach the composite generator is only
    run once at the beginning, as a setup process.

    \subsection{Structure Evolution}
    % This sentence needs more work -MG

    %During the original phase of the design of the system, it was decided that the way 
    %composites were going to evolve was by first placing only base pieces on a level 
    %then after a simulation the remaining pieces
    %of a member of the population were going to be used as a new composite if after a 
    %simulation a group was still standing and grouped on a single point, their individual 
    %offsets from the center of the group was calculated and then using this information a new 
    %composite could be added to the pool of options, this new
    % we don't use member of the populations, use proposed solution, candidate solution
    % or individual of the population. - MG  
    %additions to the composite pool could then be added to the population by the mutation
    %operation when a new member of the population was being created, the objective 
    %in this first phase of the project was to generate
    %structures that were capable of maintaining their own balance as shown on figure
    %\ref{test_old}, since the objective of open-ended evolution for this project is
    %to add new composites to the pool of options, then the algorithm will first run a
    %simulation with a group of pieces and after it the ones that were not destroyed
    %by falling because of gravity will be saved and added to the pool of options.

@@

    Structures are the next evolutionary layer, at a higher level of composition.
    Structures are compositions of composites. A level can be seen as a collection 
    of structures and each structure is a pile of composite blocks. This approach 
    was proposed by \cite{ }. This way of constructing levels has the advantage of
    limiting the search space, to structures that have a better chance to be 
    standing after the simulation ends. There is no warranty that piles of 
    composite blocks will maintain balance, so again we must search for those 
    structures.

    Structures are represented in a way similar to Togelius et al.
    \cite{togelius2016Representationsforsearch-basedmethods}. Each block has
    an id assigned to it. A structure is then a list of ids, refering to
    composite blocks, the first 11 are basic composite and generated composites 
    are assigned a higher number.  This way and individual can be
    represented as shown on figure \ref{old_chrom} where each chromosome
    represents a certain item in the list of blocks, the numbers for this
    elements are assigned in a first come first served basis, the numbers are
    infinitely incremented and the same blocks can be repeated in a structure.

@@

    These structures represent individuals in the evolutionary algorithm. Once
    an individual is created it is ready to be evaluated in a simulation on the
    game environment. Using a similar approach as the composite generator, all
    individuals have a maximum of ten seconds to maintain balance or reach
    absolute stability, if absolute stability is reached it means one of two
    things, first the generated structure is completely stable or second, the
    tower fell to either side and the pieces were destroyed or stopped moving by
    gravitational pull, when either of this happens by more than two seconds the
    simulation ends regardless of time remaining and the next individual is
    simulated.

    The evaluation process for this project is still in the development stages,
    and in our experience is one one of the more difficult components in this
    kind of systems. In this stage of the evolution, we first need to assign a
    good fitness to those structures that keep their balance and not fall apart.
    There are other variables we take in to account, but they will be explained
    later. In this initial stage, each individual in the population has 100
    points so before the simulation starts each of them are viable candidates to
    be an elite member of the population regardless of their relative height or
    complexity, then the remaining of the process is separated in three phases
    has follows. This is the same approach used in the composite generator.

@@

    Second, after removing a percentage according to the missing pieces an individual
    remaining pieces are checked for their positions in comparison to the original
    ones before the simulation, using the formula 1 each piece within the individual
    is evaluated according to the center point of the piece, if the piece moved more
    than a certain threshold half the percentage for one piece is removed from the
    remaining percentage, after this using formula 2 the position of the piece
    is further checked, this time by the difference in angle from the original one,
    in case the difference of angles is more than a certain threshold half a piece
    value is further removed from the total.
